---
title: "W10 Quiz Case Study"
author: "Valerie Ingalls"
date: "2023-03-20"
output: html_document
---

# Recreate the case study
Reference: <https://r4ds.had.co.nz/tidy-data.html#case-study
Links to an external site.>

The `who` dataset is available through the `tidyr` package as part of the `tidyverse`.
```{r}
require(tidyr)
data(who) # this will load the who dataset into memory
who # this will print out a snippet of the dataset
```

To start, we need to deal with all of the extra columns that are largely empty. We can assume that these columns actually represent values, not variables, and therefore that it would be appropriate to gather them into a column.

```{r}
who1 <- who %>%
  pivot_longer(cols = new_sp_m014:newrel_f65, 
    names_to = "key", 
    values_to = "cases", 
    values_drop_na = TRUE
  )
who1
```

We now have a "key" variable that can be meaningfully interpreted with the help of the data key; each component of the string gives information regarding the characteristics of that particular TB case. After fixing a minor error in the name, we will then separate out these components into individual columns.

```{r}
who2 <- who1 %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel")) %>%
  separate(key, c("new", "type", "sexage"), sep = "_")
who2
```
This split up the values in the code by the underscore character. We'll do another pass to remove the useless "new" column, as well as the superfluous country codes (`iso2` and `iso3`).

```{r}
who4 <- who2 %>% 
  select(-new, -iso2, -iso3)
```

We can also split up `sexage` into separate columns for sex and age

```{r}
who5 <- who4 %>% 
  separate(sexage, c("sex", "age"), sep = 1)
who5
```

This is now a tidy dataset! You can also do all of that in one big long pipe:

```{r}
tidy_who <- who %>%
  pivot_longer(
    cols = new_sp_m014:newrel_f65, 
    names_to = "key", 
    values_to = "cases", 
    values_drop_na = TRUE
  ) %>% 
  mutate(
    key = stringr::str_replace(key, "newrel", "new_rel")
  ) %>%
  separate(key, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```


## Exercises

#### 1. 

In this case study I set values_drop_na = TRUE just to make it easier to check that we had the correct values. Is this reasonable? Think about how missing values are represented in this dataset. Are there implicit missing values? What’s the difference between an NA and zero?  


Yes, this is reasonable with this dataset. NA's in this case were used when a country didn't have data for a given year. In cases where the number of TB cases for a certain category was 0, then a 0 was reported in the dataset. As NAs are in fact missing data, then, we are free to get rid of them.


#### 2. 

What happens if you neglect the mutate() step? (mutate(names_from = stringr::str_replace(key, "newrel", "new_rel")))

```{r}
test_who <- who %>%
  pivot_longer(
    cols = new_sp_m014:newrel_f65, 
    names_to = "key", 
    values_to = "cases", 
    values_drop_na = TRUE
  ) %>% 
#  mutate(
#    key = stringr::str_replace(key, "newrel", "new_rel")
#  ) %>%
  separate(key, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
test_who
```

The warning that R throws without the mutate() step tells us that in 2580 rows, there is a "missing piece". We know that the step is supposed to insert a delimiter for our 1st separate function where one is missing. With fewer delimiters than desired new columns, R is forced to fill in NAs where it doesn't perceive that is has a value to use.


#### 3. 

I claimed that iso2 and iso3 were redundant with country. Confirm this claim.

```{r}
who %>%
  select(c(country, iso2, iso3)) %>%
  distinct() %>%
  count(country) %>%
  .$n == 1
```

If we select down to only `country`, `iso2`, and `iso3`, we can then filter for distinct rows only. If there were ever any variation in what value of `iso2` or `iso3` matches up with a given value of `country`, then there would be multiple rows for that country. Count confirms that there is only one row for each function; we can use a quick boolean to print out a vector full of TRUEs for quick visual confirmation that there is only one row per country.

#### 4.

For each country, year, and sex compute the total number of cases of TB. Make an informative visualisation of the data.

```{r}
country_total <- tidy_who %>%
  group_by(country) %>%
  summarize(total = sum(cases))

year_total <- tidy_who %>%
  group_by(year) %>%
  summarize(total = sum(cases))

sex_total <- tidy_who %>%
  group_by(sex) %>%
  summarize(total = sum(cases))

country_total %>%
  ggplot(mapping = aes(x = country, y = total)) +
  geom_col()

year_total %>%
  ggplot(mapping = aes(x = year, y = total)) +
  geom_point() +
  geom_smooth(method = "loess")

sex_total %>%
  ggplot(mapping = aes(x = sex, y = total)) +
  geom_col()
```






