---
title: "Zinc"
author: "Valerie Ingalls"
date: "2022-10-28"
output: html_document
---

```{r setup, include=FALSE}
library(foreign)
library(survey)
library(tidyverse)
library(stringr)
library(bestNormalize)
library(ggplot2)
library(moments)
```

First, we have to read all of our data files into R. There are 5 different files for each cycle.

```{r import}
demo11 <- read.xport("../Data/RAW_data/DEMO_G.XPT")
auq11 <- read.xport("../Data/RAW_data/AUQ_G.XPT")
zinc11 <- read.xport("../Data/RAW_data/CUSEZN_G.XPT")
aux11 <- read.xport("../Data/RAW_data/AUX_G.XPT")
fast11 <- read.xport("../Data/RAW_data/FASTQX_G.XPT")

demo15 <- read.xport("../Data/RAW_data/DEMO_I.XPT")
auq15 <- read.xport("../Data/RAW_data/AUQ_I.XPT")
zinc15 <- read.xport("../Data/RAW_data/CUSEZN_I.XPT")
aux15 <- read.xport("../Data/RAW_data/AUX_I.XPT")
fast15 <- read.xport("../Data/RAW_data/FASTQX_I.XPT")
```

We're going to process this data all together, so we can combine our 10 dataframes into 1 here at the start. Sheets don't have matching columns within cycles so we use full join to get all of our data. Final join could be whatever variety.

```{r join}
zinc_data15 <- full_join(demo15, auq15) %>%
  full_join(zinc15) %>%
  full_join(aux15) %>%
  full_join(fast15)

zinc_data11 <- full_join(demo11, auq11) %>%
  full_join(zinc11) %>%
  full_join(aux11) %>%
  full_join(fast11)

zinc_all <- full_join(zinc_data11, zinc_data15)
```

First step in wrangling is to select only the columns that we need for the ensuing steps. This will makes a few later steps run a bit faster than if we had all columns remaining.

```{r select}
zinc_trimmed <- zinc_all %>%
  select(c(SEQN, SDDSRVYR, RIAGENDR, RIDAGEYR, RIDRETH3, WTINT2YR, WTMEC2YR, SDMVPSU, SDMVSTRA, AUQ191, AUQ280, AUQ250, AUQ260, AUQ100, AUXU500R, AUXU1K1R, AUXU2KR, AUXU500L, AUXU1K1L, AUXU2KL, AUXU3KR, AUXU4KR, AUXU6KR, AUXU8KR, AUXU3KL, AUXU4KL, AUXU6KL, AUXU8KL, AUQ330, AUQ331, AUQ370, AUQ300, WTSA2YR, LBXSCU, LBXSZN, PHDSESN))
```

Here we 
1. rename our variables to be legible 
2. refactor existing variables 
3. compute some new variables 
4. remove unwated values from the data (7 and 9 answer codes are implicit NAs, for example) 
5. trim off any remaining unwanted variables 


```{r cleaning}
# rename
zinc_refactor <- zinc_trimmed %>%
  rename("cycle" = "SDDSRVYR",
         "sex" = "RIAGENDR",
         "age_yrs" = "RIDAGEYR",
         "ethnicity" = "RIDRETH3",
         "interview_weight" = "WTINT2YR",
         "examination_weight" = "WTMEC2YR",
         "subsample_weight" = "WTSA2YR",
         "psu" = "SDMVPSU",
         "stratum" = "SDMVSTRA",
         "tinnitus_any" = AUQ191,
         "tinnitus_severity" = AUQ280,
         "tinnitus_duration" = AUQ250,
         "noise_only" = AUQ260,
         "speech_in_noise" = AUQ100,
         "non_work_noise" = AUQ370,
         "firearms" = AUQ300,
         "copper" = LBXSCU,
         "zinc" = LBXSZN,
         "session" = PHDSESN
  ) %>%
# refactor, compute, combine for various variables
  mutate(
         "work_noise" = if_else(!is.na(AUQ330), AUQ330, AUQ331), # different question name between cycles, just a combine + rename here
         "ethnicity" = if_else(`ethnicity` == 3, 1, 2), # 1 = white, 2 is non white/Hispanic
         "age_group" = if_else(age_yrs < 30, 1, # youngest individuals with audio data are 20
                               if_else(age_yrs < 40, 2, # so age ranges are 20-29, 30-39, etc
                                       if_else(age_yrs < 50, 3, 
                                               if_else(age_yrs < 60, 4, 5)))), # upper limit was 69
         across(c(AUXU500R, AUXU1K1R, AUXU2KR, AUXU3KR, AUXU4KR, AUXU6KR, AUXU8KR, AUXU500L, # removing null hearing thresholds
                   AUXU1K1L, AUXU2KL, AUXU3KL, AUXU4KL, AUXU6KL, AUXU8KL), ~na_if(., 888)),
         across(c(AUXU500R, AUXU1K1R, AUXU2KR, AUXU3KR, AUXU4KR, AUXU6KR, AUXU8KR, AUXU500L,
                   AUXU1K1L, AUXU2KL, AUXU3KL, AUXU4KL, AUXU6KL, AUXU8KL), ~if_else(. == 666, 110, .)), # thresholds above measured values are being set to 110 (highest attempted measure)
         "mid_pta_right" = (AUXU500R + AUXU1K1R + AUXU2KR) / 3, # standard Pure Tone Average calc
         "mid_pta_left" = (AUXU500L + AUXU1K1L + AUXU2KL) / 3,
         "high_pta_right" = (AUXU3KR + AUXU4KR + AUXU6KR + AUXU8KR) / 4, # same but for higher freq
         "high_pta_left" = (AUXU3KL + AUXU4KL + AUXU6KL + AUXU8KL) / 4,
         "combined_subsample_weight" = subsample_weight/2, # proper method for weight when using multiple cycles. Take weight, divide by number of cycles
         across(c(tinnitus_any, tinnitus_severity, tinnitus_duration, noise_only, sex, non_work_noise, firearms, work_noise, speech_in_noise),  ~na_if(., 7)), # na coded null values
         across(c(tinnitus_any, tinnitus_severity, tinnitus_duration, noise_only, sex, non_work_noise, firearms, work_noise, speech_in_noise),  ~na_if(., 9)),
         "problematic" = if_else(tinnitus_severity == 1, 0, 1), # tinnitus that is at least a small problem vs not
         "acute_chronic" = if_else(tinnitus_duration >= 3, 1, 0), # tinnitus lasting at least a year vs not
         "tinnitus_type" = if_else(tinnitus_any == 2, 3, 1),
         "tinnitus_type" = if_else(!is.na(problematic) & problematic == 0, 2, 
                                   if_else(!is.na(acute_chronic) & acute_chronic == 0, 2,
                                           if_else(!is.na(noise_only) & noise_only == 1, 2, tinnitus_type))),
# 1: bothersome chronic, 2: not bothersome or not chronic or noise-exposure only, 3: no tinnitus
         "cycle" = if_else(cycle == 7, "2011-2012", "2015-2016"), # changing cycle code to a year for clarity
         "zinc_cutoff" = if_else(age_yrs >= 10, # using standards to control zinc deficiency cutoff based on known factors. no such info found on copper, so that cutoff is less controlled
                                     if_else(sex == 1, if_else(session == 1, 74, 61), 
                                     if_else(session == 1, 70, 59)),
                                     if_else(session == 1, 65, 57)),
         "zinc_deficiency" = if_else(`zinc` < zinc_cutoff, 1, 0), # 0 is no deficiency
         "copper_deficiency" = if_else(`copper` < 63.7, 1, 0),
         "tinnitus_any" = if_else(tinnitus_any == 1, 1, 0)
  ) %>%
# remove remaining unwanted variables
  select(-c(AUXU500R, AUXU1K1R, AUXU2KR, AUXU500L, AUXU1K1L, AUXU2KL, AUXU3KR, AUXU4KR, AUXU6KR, AUXU8KR, AUXU3KL, AUXU4KL, AUXU6KL, AUXU8KL, AUQ330, AUQ331, subsample_weight, zinc_cutoff, session, problematic, acute_chronic, noise_only))
    
```




```{r subset}
# file without NA's for easier use with SPSS; no purpose here
#zinc_SPSS <- zinc_refactor %>%
#    replace(is.na(.), "")


# removing rows with incomplete data; partially for cleanliness, partially for better functioning with later analyses
zinc_subset <- zinc_refactor %>%
  filter(!is.na(zinc)) %>%
  filter(!is.na(copper)) %>%
  filter(!is.na(tinnitus_any)) %>%
  filter(!is.na(speech_in_noise)) %>%
  filter(!is.na(tinnitus_type))
```


In preliminary analyses, it was noted that there is skew in the values of serum copper and zinc. Therefore, we will perform a transformation on the data to attempt to create more normal distribution.
```{r normalization}

zinc_norm_result <- bestNormalize(zinc_subset$zinc)
normed_zinc <- zinc_norm_result$x.t


copper_norm_result <- bestNormalize(zinc_subset$copper)
normed_copper <- copper_norm_result$x.t



skewness(zinc_subset$zinc)
skewness(normed_zinc)


skewness(zinc_subset$copper)
skewness(normed_copper)

```

In both cases we can see that these methods have reduced the skewness of the distribution, so we will attempt analyses using the normalized zinc and copper distributions.

```{r appendNorm}
zinc_final <- zinc_subset %>%
  mutate("normedZinc" = normed_zinc,
         "normedCopper" = normed_copper)
```




```{r write}
# final output csv that we're going to read into our analysis rmd
write.csv(zinc_final, file = "../Data/processed_metal_data.csv")
```
