---
title: "Zinc"
author: "Valerie Ingalls"
date: "2022-10-28"
output: html_document
---

```{r setup, include=FALSE}
library(survey)
library(tidyverse)
library(ggplot2)
```


```{r import}
# grabbing our finalized data sheet from generated from our processing script earlier
df <- read.csv("../Data/processed_metal_data.csv")
```


```{r factorize}
# setting a var to factor and removing the csv artifact column
df$tinnitus_type <- as.factor(df$tinnitus_type)
levels(df$tinnitus_type) <- c("Chronic Bothersome Tinnitus", "Other Tinnitus", "No Tinnitus")

df <- select(df, -X)
```


We next establish our design variable, which is part of the `survey` package. This will allow us to include the elements of the complex survey design in our statistical analyses for better generalization to the population level.
```{r design}
tinnitus_design <- svydesign(data = df, id = ~psu, strata = ~stratum, weights = ~combined_subsample_weight, nest = TRUE)
```

First, chiSquare tests to look at the differences in demographic variables across our tinnitus groups.
```{r chisq}
weighted_sextest <- svychisq(~tinnitus_type + sex, design = tinnitus_design)
weighted_sextest$observed
weighted_sextest


weighted_ethtest <- svychisq(~tinnitus_type + ethnicity, design = tinnitus_design)
weighted_ethtest$observed
weighted_ethtest
```

Multiple linear regression models here. Running separate models with zinc and copper as outcome variables and using transformed  values. 
```{r multiple_lin_reg}
lmzn <- svyglm(normedZinc ~ tinnitus_type + age_yrs + ethnicity + sex + speech_in_noise, design = tinnitus_design)
summary(lmzn)


lmcu <- svyglm(normedCopper ~ tinnitus_type + age_yrs + ethnicity + sex + speech_in_noise, design = tinnitus_design)
summary(lmcu)

```


next is logistic regression to examine deficiency of each of the metals.
```{r}
zn_def_logit <- svyglm(zinc_deficiency ~ tinnitus_type + age_yrs + ethnicity + sex + speech_in_noise, design = tinnitus_design, family = "quasibinomial")
summary(zn_def_logit)


cu_def_logit <- svyglm(copper_deficiency ~ tinnitus_type + age_yrs + ethnicity + sex + speech_in_noise, design = tinnitus_design, family = "quasibinomial")
summary(cu_def_logit)

```


While this is not going to be the most useful visualization due to the low difference between these groups, we will still begin with our original plan of visualizing the difference between metal levels in tinnitus groups via a nested density plot. Weights are incorporated into the plots to better account for survey design.
```{r histograms}
# non-normalized density plots

ggplot(data = df) +
  geom_density(mapping = aes(x = zinc, weight = combined_subsample_weight, fill = tinnitus_type), alpha = .5) +
  geom_vline(xintercept = 61, color = "blue", linetype = 2) +
  labs(x = "Serum Zinc Level (\u03BCg/dL)", y = "Density", title = "Distribution of Serum Zinc Levels by Tinnitus Type", fill = "Tinnitus Type")
#ggsave("../Output/Figures/Final_Zinc_Density.png")


ggplot(data = df) +
  geom_density(mapping = aes(x = copper, weight = combined_subsample_weight, fill = tinnitus_type), alpha = .5) +
  geom_vline(xintercept = 63.7, color = "red", linetype = 2) +
  labs(x = "Serum Copper Level (\u03BCg/dL)", y = "Density", title = "Distribution of Serum Copper Levels by Tinnitus Type", fill = "Tinnitus Type")
#ggsave("../Output/Figures/Final_Copper_Density.png")

```

To show just how similar these distributions are, we're going to also create box plots. The `svyboxplot()` function from `survey` allows us to do this while fully implementing our sample design.
```{r}
# adjusts text spacing so it all fits
par(mar = c(5, 5, 4, 2) + .1)

svyboxplot(normedZinc~tinnitus_type, design = tinnitus_design, xlab = "Tinnitus Type", ylab = expression("Serum Zinc Level ( " * sqrt(mu *g/dL) * ")"), main = "Serum Zinc Levels by Tinnitus Type", all.outliers = TRUE, col = "#00beff", boxwex = .75)



par(mar = c(5, 5, 4, 2) + .1, cex.lab = .78)

svyboxplot(normedCopper~tinnitus_type, design = tinnitus_design, all.outliers = TRUE, xlab = "Tinnitus Type", ylab = expression("Serum Copper Level (Yeo-Johnson Transformed " * mu *"g/dL)"), main = "Serum Copper Levels by Tinnitus Type", col = "#00beff", boxwex = .75)

```
